MATH/COMP 562 ASSIGNMENT 2

### 2023-06-29 assignment-2/article.tex
Solve 1.2
Solve 1.3

\section{Convex Learning Problems}
Refer to Ch 12 of Understanding Machine Learning (Shalev-Shwartz).

>>>>

\begin{solution}
  Let $S$ and $S^\prime$ be two datasets differing in one point, say $z_m\in S$ and $z_m^\prime \in S^\prime$. We proceed by bounding the expectation $\hat{\mathbb{E}}_S[\Phi(S)]$ as follows:
  \begin{align*}
    \hat{\mathbb{E}}_S[\Phi(S)] & = \hat{\mathbb{E}}_S\left[ \sup_{h\in\mathcal{H}} \left| \mathbb{E}[h] - \hat{\mathbb{E}}_S[h] \right| \right]                                                                                       \\
                                & = \mathbb{E}_S \left[ \sup_{h\in\mathcal{H}} \hat{\mathbb{E}}_{S^\prime}[h] - \hat{\mathbb{E}}_{S}[h] \right]                                                                                        \\
                                & \leq \mathbb{E}_{S,S^\prime}\left[ \sup_{h\in\mathcal{H}} \hat{\mathbb{E}}_{S^\prime}[h] - \hat{\mathbb{E}}_{S}[h] \right]                                                                           \\
                                & = \mathbb{E}_{S,S^\prime} \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m\left(h(z_i^\prime) - h(z_i)\right) \right]                                                                          \\
                                & = \mathbb{E}_{\sigma, S, S^\prime} \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (h(z_i^\prime) - h(z_i)) \right]                                                                  \\
                                & \leq \mathbb{E}_{\sigma, S^\prime}\left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sigma_i h(z_i^\prime) \right] + \mathbb{E}_{\sigma, S}\left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sigma_i h(z_i) \right] \\
                                & = 2\mathbb{E}_{\sigma,S}\left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma h(z_i) \right] = 2\hat{\mathfrak{R}}_m(\mathcal{H})
  \end{align*}
\end{solution}

\begin{problem}{Exercise 1.3 (Rademacher Identities)}
Fix $m\geq 1$. Prove the following identities for any $\alpha \in \mathbb{R}$ and any two hypothesis sets $\mathcal{H}$ and $\mathcal{H^\prime}$ of functions mapping from $\mathcal{X}$ to $\mathcal{R}$.
\begin{enumerate}[label= (\alph*)]
  \item $\mathfrak{R}_m(\alpha \mathcal{H}) = |\alpha| \mathfrak{R}_m(\mathcal{H})$.
  \item $\mathfrak{R}_m(\mathcal{H} + \mathcal{H}^\prime) = \mathfrak{R}_m(\mathcal{H}) + \mathfrak{R}_m(\mathcal{H^\prime})$.
\end{enumerate}
\end{problem}

\begin{solution}
  \begin{enumerate}[label= (\alph*)]
    \item We can bound $\mathfrak{R}_m(\alpha \mathcal{H})$ above and below and establish our result using the squeeze theorem.
          \begin{align*}
            \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (-|\alpha| h(z_i)) \right]                & \leq \mathfrak{R}_m(\alpha \mathcal{H}) & \leq \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (|\alpha| h(z_i)) \right] \\
            \mathbb{E}_{\sigma^\prime} \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i^\prime (|\alpha| h(z_i)) \right] & \leq \mathfrak{R}_m(\alpha \mathcal{H}) & \leq \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (|\alpha| h(z_i)) \right]
          \end{align*}
          Both sides of the last inequality are equal to each other since $\sigma$ and $\sigma^\prime = -\sigma$ are identically distributed. Lastly we use linearity of expectation and the independence of $\alpha$ on $h$ to shift the $|\alpha|$ term outside of the expectation in the following way:
          \begin{align*}
            \mathfrak{R}_m(\alpha \mathcal{H}) & = \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (|\alpha| h(z_i)) \right] & = |\alpha| \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i h(z_i) \right] \\
                                               & = |\alpha| \mathfrak{R}_m(\mathcal{H}) \text{ this concludes the proof}
          \end{align*}
          \newpage
    \item We will rely on the fact that $h$ and $h^\prime$ do not depend on one another so maximising the sum is the same as maximising each separately. As a result, we can use the following argument: \begin{align*}
            \mathfrak{R}_m(\mathcal{H} + \mathcal{H^\prime}) & = \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}, h^\prime \in \mathcal{H}^\prime} \frac{1}{m} \sum_{i=1}^m \sigma_i (h(z_i) + h^\prime (z_i)) \right]                                                                           \\
                                                             & = \mathbb{E}_\sigma \left[ \sup_{h\in\mathcal{H}} \frac{1}{m} \sum_{i=1}^m \sigma_i (h(z_i)) \right] + \mathbb{E}_\sigma \left[ \sup_{h^\prime \in \mathcal{H}^\prime} \frac{1}{m} \sum_{i=1}^m \sigma_i (h^\prime(z_i)) \right] \\
                                                             & = \mathfrak{R}_m(\mathcal{H}) + \mathfrak{R}_m(\mathcal{H}^\prime)
          \end{align*}

  \end{enumerate}
\end{solution}
\section{Convex Learning Problems}
Refer to Ch 12 of Understanding Machine Learning (Shalev-Shwartz).
